### 2016-08-16 cover_letter/cover.tex
discussion, cover letter, minor tweaks to text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Professional Formal Letter
% LaTeX Template
% Version 1.0 (28/12/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Brian Moses (http://www.ms.uky.edu/~math/Resources/Templates/LaTeX/)
% with extensive modifications by Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------------------------
%   PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%------------------------------------------------------------------------------

\documentclass[11pt,a4paper]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)

\usepackage{graphicx} % Required for including pictures
\usepackage{microtype} % Improves typography
\usepackage[T1]{fontenc} % Required for accented characters
\usepackage[utf8]{inputenc}

% Create a new command for the horizontal rule in the document which allows thickness specification
\makeatletter
\def\vhrulefill#1{\leavevmode\leaders\hrule\@height#1\hfill \kern\z@}
\makeatother

%----------------------------------------------------------------------------------------
%   DOCUMENT MARGINS
%----------------------------------------------------------------------------------------

\textwidth 6.75in
\textheight 9.25in
\oddsidemargin -.25in
\evensidemargin -.25in
\topmargin -1in
\parindent 0.4in

%------------------------------------------------------------------------------
%   SENDER INFORMATION
%------------------------------------------------------------------------------

\def\Who{Eduard Szöcs} % Your name
\def\Where{University of Koblenz-Landau\\ Institute for Environmental Sciences} % Your department/institution
\def\Address{Fortstr. 7} % Your address
\def\CityZip{76829 Landau - GERMANY} % Your city, zip code, country, etc
\def\Email{E-mail: szoecs@uni-landau.de} % Your email address
\def\TEL{Phone: +49 6321 280 31552} % Your phone number

%------------------------------------------------------------------------------
%   HEADER AND FROM ADDRESS STRUCTURE
%------------------------------------------------------------------------------

\address{
    \hspace{\fill}\includegraphics[width=3in]{Uni-Logo-2c.jpg} % Include the logo 
    ~\\[-0.11in] % Reduce the whitespace above the horizontal rule
    \vhrulefill{1pt} \\ 
    \hspace{\fill}\parbox[t]{4in}{ % Create a box for your details underneath the horizontal rule on the right
        \vspace{0.6in}
        \textbf{\Who} \em \\ % Your name, all text after this will be italicized
        \Where\\ % Your department
        \Address\\ % Your address
        \CityZip\\ % Your city and zip code
        \TEL\\ % Your phone number
        \Email\\ % Your email address
    }
    \hspace{-1.4in} % Horizontal position of this block, increase to move left, decrease to move right
    \vspace{-1in} % Move the letter content up for a more compact look
}

%------------------------------------------------------------------------------
%   TO ADDRESS STRUCTURE
%------------------------------------------------------------------------------

\def\opening#1{
\thispagestyle{empty}
{\centering\fromaddress \vspace{1in} \\} 
{\raggedright \toname \\ \toaddress \par} % Print the to name and address
\vspace{2em}
% Subject line
\noindent
\textbf{Subject line} \hspace*{\fill}\today \\[2em]
\vspace{0.6in} % White space after the to address
\noindent #1 % Print the opening line
}

%----------------------------------------------------------------------------------------
%   SIGNATURE STRUCTURE
%----------------------------------------------------------------------------------------

\signature{\Who} % The signature is a combination of your name and title

\long\def\closing#1{
    \vspace{0.1in} % Some whitespace after the letter content and before the signature
    \noindent % Stop paragraph indentation
    \hspace*{\fill} % Move the signature right
    \parbox{\indentedwidth}{\raggedright
    #1 % Print the signature text
    \vskip 0.65in % Whitespace between the signature text and your name
    \fromsig}
} % Print your name and title

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%   TO ADDRESS
%----------------------------------------------------------------------------------------

\begin{letter}
{
\textbf{Environmental Science \& Technology} \\
\vspace{1em}
To the editor
}

%------------------------------------------------------------------------------
%   LETTER CONTENT
%------------------------------------------------------------------------------

\opening{Dear editor of Environmental Science \& Technology,}

xxxxx

\closing{Sincerely, }

%----------------------------------------------------------------------------------------

\end{letter}
\end{document}