### 2016-08-19 cover_letter/cover.tex
cover letter + refs

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Professional Formal Letter
% LaTeX Template
% Version 1.0 (28/12/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Brian Moses (http://www.ms.uky.edu/~math/Resources/Templates/LaTeX/)
% with extensive modifications by Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------------------------
%   PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%------------------------------------------------------------------------------

\documentclass[11pt,a4paper]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)

\usepackage{graphicx} % Required for including pictures
\usepackage[T1]{fontenc} % Required for accented characters
\usepackage[utf8]{inputenc}

% Create a new command for the horizontal rule in the document which allows thickness specification
\makeatletter
\def\vhrulefill#1{\leavevmode\leaders\hrule\@height#1\hfill \kern\z@}
\makeatother

%----------------------------------------------------------------------------------------
%   DOCUMENT MARGINS
%----------------------------------------------------------------------------------------

\textwidth 6.75in
\textheight 9.25in
\oddsidemargin -.25in
\evensidemargin -.25in
\topmargin -1.5in
\parindent 0in

%------------------------------------------------------------------------------
%   SENDER INFORMATION
%------------------------------------------------------------------------------

\def\Who{Eduard Szöcs} % Your name
\def\Where{University of Koblenz-Landau\\ Institute for Environmental Sciences} % Your department/institution
\def\Address{Fortstr. 7} % Your address
\def\CityZip{76829 Landau - GERMANY} % Your city, zip code, country, etc
\def\Email{E-mail: szoecs@uni-landau.de} % Your email address
\def\TEL{Phone: +49 6321 280 31552} % Your phone number

%------------------------------------------------------------------------------
%   HEADER AND FROM ADDRESS STRUCTURE
%------------------------------------------------------------------------------

\address{
    % Include the logos
    \includegraphics[width=3in]{Uni-Logo-2c.jpg}\hspace{\fill}\includegraphics[width=2in]{logo_institut.jpg}
    ~\\[-0.11in] % Reduce the whitespace above the horizontal rule
    \vhrulefill{1pt} \\ 
    \hspace{\fill}\parbox[t]{4in}{ % Create a box for your details underneath the horizontal rule on the right
        \vspace{0.2in}
        \textbf{\Who} \em \\ % Your name, all text after this will be italicized
        \Where\\ % Your department
        \Address\\ % Your address
        \CityZip\\ % Your city and zip code
        \TEL\\ % Your phone number
        \Email\\ % Your email address
    }
    \hspace{-1.4in} % Horizontal position of this block, increase to move left, decrease to move right
    \vspace{-1in} % Move the letter content up for a more compact look
}

%------------------------------------------------------------------------------
%   TO ADDRESS STRUCTURE
%------------------------------------------------------------------------------

\def\opening#1{
\thispagestyle{empty}
{\centering\fromaddress \vspace{0.5in} \\} 
{\raggedright \toname \\ \toaddress \par} % Print the to name and address
\vspace{2em}
% Subject line
\noindent
\textbf{Submission of article} \hspace*{\fill}\today \\[2em]
\vspace{0.6in} % White space after the to address
\noindent #1 % Print the opening line
}

%----------------------------------------------------------------------------------------
%   SIGNATURE STRUCTURE
%----------------------------------------------------------------------------------------

\signature{\Who} % The signature is a combination of your name and title

\long\def\closing#1{
    \vspace{0.1in} % Some whitespace after the letter content and before the signature
    \noindent % Stop paragraph indentation
    \hspace*{\fill} % Move the signature right
    \parbox{\indentedwidth}{\raggedright
    #1 % Print the signature text
    \vskip 0.65in % Whitespace between the signature text and your name
    \fromsig}
} % Print your name and title

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%   TO ADDRESS
%----------------------------------------------------------------------------------------

\begin{letter}
{
\textbf{Environmental Science \& Technology} \\
\vspace{1em}
To the editor
}

% noindent
\setlength\parindent{0pt}

%------------------------------------------------------------------------------
%   LETTER CONTENT
%------------------------------------------------------------------------------

\opening{Dear Editor,}
\vspace{-1em}\\
enclosed is a manuscript to be considered for publication in Environmental Science \& Technology titled \\

\emph{Pesticides in small agricultural streams in Germany}\\

Pesticides enter freshwater ecosystems from agricultural areas and may have effects on biota.
However, relatively little is know about their large-scale distribution.
Especially, small streams may be at high risk because of adjacency to fields and small water columns.

We compiled and homogenized monitoring data from all German Federal States with a focus on small to mid-size streams. 
Our results show that there is lack of knowledge for small streams. 
However, the data revealed the pesticides are commonly occurring.
Especially, neonicotinoid pesticides may pose a risk to small agricultural freshwaters.

Our study should be of general interest and add valuable information on chemical monitoring, pesticide occurrences and ecological risk arising from them.

If you have any further questions do not hesitate to contact me.

\closing{Yours sincerely, }

%----------------------------------------------------------------------------------------

\end{letter}
\end{document}