### 2016-10-18 supplement/Supplemental_Materials.tex
more revisions

\documentclass[pdftex,a4paper]{scrreprt}
\usepackage[utf8]{inputenc}       
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{booktabs} % table rulers
\usepackage{longtable} % multipage tables
\usepackage{todonotes}
\usepackage{amsmath}
\usepackage{float} % for [H]

% Continuous figure counting (instead of per cahpter)
\usepackage{chngcntr}
\counterwithout{figure}{chapter}
\counterwithout{table}{chapter}
% Add 'S' before figure/table numbering
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

\usepackage[
	round,	%(defaultage in the main file and \input ) for round parentheses;
	authoryear,% (default) for author-year citations;
	sort,		% orders multiple citations into the sequence in which they 
]{natbib}

%% -------------------------------------------------------------------------
\title{Supplemental Materials for the paper: \\ Pesticide exposure in small streams in Germany}
\author{Eduard Szöcs, Marvin Brinke, Bilgin Karaoglan, Ralf B. Schäfer}

%% -------------------------------------------------------------------------
\begin{document}
\maketitle

\tableofcontents
\listoffigures
\listoftables

%% -------------------------------------------------------------------------
\chapter{Data Cleaning}
More then 30 datasets have been cleaned and homogenized separately, before combing into a common database.
Cleaning steps comprised (Figure~\ref{fig:data_cleaning} gives a graphical overview).

\begin{enumerate}
	\item Structure: Structure has been adjusted to the database structure.
	\item Coordinates: Coordinates have been transformed to a common Coordinate Reference System (DHDN / 3-Grad Gauss-Krüger Zone 3 (EPSG:31467)) and duplicates merged.
	\item Chemicals: Chemical names and identifiers have been unified using the webchem package \citep{szocs_webchem:_2016}.
	\item  Identifiers: Unique identifiers have been assigned.
	\item Units: All concentrations have been converted to $\mu g/L$. Values below limit of quantification have be set to zero.
	\item Other meta-data: meta-data has been standardised.
	\item Temporal resolution: The temporal resolution of the database is 1 day. Date below this resolution has been aggregated by maximum.
	\item Validity Checks: Simple rules for validity checks have been implemented (e.g. no negative concentrations).
\end{enumerate}

\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{data_cleaning}
	\caption[Overview on data cleaning steps.]{Overview on data cleaning steps. After cleaning data has been stored in a relational spatial PostgreSQL database.}
	\label{fig:data_cleaning}
\end{figure}

%% -------------------------------------------------------------------------
\chapter{Overview on compiled data}
% Overview samples
% table generated in do_overview.R
\input{phchoverview}

\begin{figure}[ht]
	\centering
	\includegraphics[width = \textwidth]{temporal}
	\caption{Number of sampling occasions per year and month.}
	\label{fig:temporal}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{varclus}
	\caption[Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states.]{Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states. Abbreviations of state names according to ISO 3166-2:DE.}
	\label{fig:varclus}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.6\textwidth]{silhouette}
	\caption[Average silhouette width for different cluster sizes.]{Average silhouette width for different cluster sizes. Two clusters showed maximum silhouette width.}
	\label{fig:silhouette}
\end{figure}

% Overview variables
% table generated in do_overview.R
\input{phchvar}

%% -------------------------------------------------------------------------
\chapter{Thresholds for agricultural land use and catchment size}
\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{ezgagrirac}
	\caption[Raw data used for the model in equation 2 and Figure 3 of the main article.]{Raw data used for the model in equation 2 and Figure 3 of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points denote sites without any exceedance.}
	\label{fig:ezgagrirac}
\end{figure}

%% -------------------------------------------------------------------------
\chapter{Effect of precipitation and season on RQ}
\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{precip}
	\caption[Distribution of precipitation at sampling occasions.]{Distribution of precipitation at sampling occasions. top: at sampling date. bottom: at day before sampling.}
	\label{fig:precip}
\end{figure}

\input{keeptab}
\input{keeptab2}

\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{coefs}
	\caption[Graphical representation of coefficients from table \ref{tab:var_model_coef}.]{Graphical representation of coefficients from table \ref{tab:var_model_coef}. Top row: Effect of precipitation at day before sampling and at day of sampling. Bottom row: estimates for the four Quarters. Each dot represent one compound (in the order described in table \ref{tab:var_model}). Coefficients where the CI encompasses zero are shown in gray colour. Coefficients are shown on the link scale (log for $\mu$ and logit for $\nu$).}
	\label{fig:coefs}
\end{figure}

%% -------------------------------------------------------------------------
\chapter{Pesticides in small streams}

\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{prac_ex}
	\caption[Cumulative distribution of the number sites exceeding RAC.]{Cumulative distribution of sites exceeding RAC. Dotted lines indicate fraction of sites exceeding a RQ of 1 and 0.1. 23\% of sites showed no detection of compounds with RAC and are not shown due to logarithmic x-axis.}
	\label{fig:pdetects}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{pmix}
	\caption{Distribution the number of quantified compounds in the samples from small streams.}
	\label{fig:pmix}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{pdetects}
	\caption[Proportion of samples with detects in small streams.]{Proportion of samples with detects in small streams. Only Compounds with more than 100 samples and 15\% of detects are show.}
	\label{fig:pdetects}
\end{figure}

\bibliographystyle{apalike}      % basic style, author-year citations
\bibliography{references}

\end{document}