### 2017-02-08 supplement/Supplemental_Materials.tex

\documentclass[pdftex,
	a4paper,
	titlepage=false]{scrreprt}
\usepackage[utf8]{inputenc}       
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{booktabs} % table rulers
\usepackage{longtable} % multipage tables
\usepackage{todonotes}
\usepackage{amsmath}
\usepackage{float} % for [H]
\usepackage{array, ragged2e}
\newcolumntype{R}[1]{>{\RaggedLeft\arraybackslash}p{#1}} % right aling column with fixed width

% Continuous figure counting (instead of per cahpter)
\usepackage{chngcntr}
\counterwithout{figure}{chapter}
\counterwithout{table}{chapter}
% Add 'S' before figure/table numbering
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thepage}{S\arabic{page}}

\usepackage[
	round,	%(defaultage in the main file and \input ) for round parentheses;
	authoryear,% (default) for author-year citations;
	sort,		% orders multiple citations into the sequence in which they 
]{natbib}

% affiliations
\usepackage[affil-it]{authblk}

% counts
% http://ctan.mackichan.com/macros/latex/contrib/caption/totalcount.pdf
\usepackage[figure,table, page]{totalcount}

%% -------------------------------------------------------------------------
\title{Supplemental Materials for: \\ Large scale risks from pesticides in small streams}
\author{Eduard Szöcs%
  \thanks{Corresponding author, Email: \texttt{szoecs@uni-landau.de}}}
\affil[1]{Institute for Environmental Sciences, University of Koblenz-Landau, Germany}

\author{Marvin Brinke}
\affil{German Federal Institute of Hydrology (BfG), Koblenz, Germany}

\author{Bilgin Karaoglan}
\affil{German Environment Agency (UBA), Dessau-Roßlau, Germany}

\author[1]{Ralf B. Schäfer}

%% -------------------------------------------------------------------------
\begin{document}
\maketitle

\vfill
\begin{description}
	\item[Number of pages: ] \totalpages
	\item[Number of figures: ] \totalfigures
	\item[Number of tables: ] \totaltables
\end{description}
% increase page counter by 1
\setcounter{page}{2}

\tableofcontents
\listoffigures
\listoftables

%% -------------------------------------------------------------------------
\chapter{Data Cleaning}
Before combining into a common database, more than 30 datasets have been cleaned and homogenised separately.
Cleaning steps comprised the following steps (Figure~\ref{fig:data_cleaning} gives a graphical overview):

\begin{enumerate}
	\item Structure: Datasets have been adjusted to the database structure.
	\item Coordinates: Coordinates have been transformed to a common Coordinate Reference System (DHDN / 3-Grad Gauss-Krüger Zone 3 (EPSG:31467)) and duplicates merged.
	\item Chemicals: Chemical names and identifiers have been unified using the webchem package (https://github.com/ropensci/webchem).
	\item  Identifiers: Unique identifiers have been assigned.
	\item Units: All concentrations have been converted to $\mu g/L$. Values below limit of quantification were set to zero (and can be used to identify non-detects).
	\item Other meta-data: meta-data has been standardised.
	\item Temporal resolution: The temporal resolution of the database is 1 day. Samplings below this resolution have been aggregated by the maximum daily value.
	\item Validity Checks: Simple rules for validity checks have been implemented.
\end{enumerate}

\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{data_cleaning}
	\caption[Overview on data cleaning steps.]{Overview on data cleaning steps. After cleaning, data have been stored in a relational spatial PostgreSQL database.}
	\label{fig:data_cleaning}
\end{figure}

%% -------------------------------------------------------------------------
\chapter{Overview on compiled data}
% Overview samples
% table generated in do_overview.R
\input{phchoverview}

\begin{figure}[ht]
	\centering
	\includegraphics[width = \textwidth]{temporal}
	\caption{Number of sampling occasions per year and month.}
	\label{fig:temporal}
\end{figure}

% \begin{figure}[ht]
% 	\centering
% 	\includegraphics[width = 0.8\textwidth]{varclus}
% 	\caption[Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states.]{Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states. Abbreviations of state names according to ISO 3166-2:DE (see also Table~\ref{tab:phch_overview}).}
% 	\label{fig:varclus}
% \end{figure}

% \begin{figure}[ht]
% 	\centering
% 	\includegraphics[width = 0.6\textwidth]{silhouette}
% 	\caption[Average silhouette width for different cluster sizes.]{Average silhouette width for different cluster sizes of complete linkage clustering of jaccard similarity of analysed compound spectra between federal states. Two clusters showed the maximum silhouette width.}
% 	\label{fig:silhouette}
% \end{figure}

% Overview variables
% table generated in do_overview.R
\newpage
\input{phchvar}

%% -------------------------------------------------------------------------
\chapter{Thresholds for agricultural land use and catchment size}
\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{ezgagrirac}
	\caption[Raw data used for the model in equation 2 and Figure 3 of the main article.]{Raw data used for the model in equation 2 and Figure 3 of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points denote sites without any exceedance.}
	\label{fig:ezgagrirac}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width = \textwidth]{smooth_lin.pdf}
	\caption[Smooth and linear fit]{Smooth and linear fits for the data presented in figure 3 of the main article.}
	\label{fig:smooth_lin}
\end{figure}

%% -------------------------------------------------------------------------
\chapter{Effect of precipitation and season on RQ}
% \begin{figure}[ht]
% 	\centering
% 	\includegraphics[width = 0.8\textwidth]{precip}
% 	\caption[Distribution of precipitation at sampling occasions.]{Distribution of precipitation at sampling occasions. top: at sampling date. bottom: at the day before sampling.}
% 	\label{fig:precip}
% \end{figure}

%%% from do_precip.R
\newpage
\input{keeptab}
\newpage
\input{keeptab2}

\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{coefs}
	\caption[Graphical representation of coefficients from table \ref{tab:var_model_coef}.]{Graphical representation of coefficients from table \ref{tab:var_model_coef}. Top row: Effect of precipitation at the day before sampling and at day of sampling. Bottom row: estimates for the four Quarters. Each dot represents one compound (in the order described in table \ref{tab:var_model}). Coefficients where the CI encompasses zero are shown in gray colour. Coefficients are shown on the link scale (log for $\mu$ and logit for $\nu$).}
	\label{fig:coefs}
\end{figure}

\begin{figure}[ht]
\centering
  \includegraphics[width=0.8\textwidth]{mean_coef.pdf}
  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter (bottom row) from a meta-analysis of the 22 modelled compounds (see Figure S6). Left: coefficients for mean RQ ($\mu$), right: coefficients for probability of exceeding LOQ ($\nu$). 
  Coefficients are shown on the link scale.
  Single compound coefficients are provided in Supplemental Table~S4 and Figure~S7).
  }
  \label{fig:mean_coefs}
\end{figure}

%% -------------------------------------------------------------------------
\chapter{Pesticides in small streams}
%% from do_pollution.R
\input{racdat}

\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{prac_ex}
	\caption[Cumulative distribution of the number sites exceeding RAC.]{Cumulative distribution of sites exceeding RAC. Dotted lines indicate fraction of sites exceeding a RQ of 1 and 0.1. 23\% of the sites showed no detection of compounds with available RAC values and are not shown due to logarithmic x-axis.}
	\label{fig:prac_ex}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{pdetects}
	\caption[Proportion of samples with detects in small streams.]{Proportion of samples with detects in small streams. Only Compounds with more than 100 samples and 15\% of detects are shown.}
	\label{fig:pdetects}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width = 0.8\textwidth]{pmix}
	\caption{Distribution of the number of quantified compounds in the samples.}
	\label{fig:pmix}
\end{figure}

% \bibliographystyle{apalike}      % basic style, author-year citations
% \bibliography{references}

\end{document}